# Statistical Learning

```{r}
library(ISLR)
library(tidyverse)
library(magrittr)
```

## Explore the College Dataset [8]

```{r}
college <- College %>%
  as_tibble() 
college
```

Skip rownames.

```{r}
college %>% summary()
```

In how many Universities more than half of the students come from the top 10% colleges?

```{r}
college <- college %>%
  mutate(elite = case_when(Top10perc > 50 ~ "Yes",
                           TRUE ~ "No"))

college %$% table(elite)
```

And are those college outstate?

```{r}
p <- college %>%
  ggplot(aes(x = elite,
             y = Outstate))

# With boxplot
p1 <- p +
  geom_boxplot(outlier.alpha = 0)
p1


# with beeswarm
p1 + ggbeeswarm::geom_quasirandom(alpha = .4)
```

Produce multiple histograms:

```{r}
college %>%
  select_if(is.numeric) %>%
  gather() %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(facet = "key",
             scales = "free", 
             ncol = 3) +
  theme_bw()
```

Report for EDA:

```{r}

```

## Explore the Auto dataset [9]

Take Auto dataset and remove missing values

```{r}
Auto %>% anyNA()

# otherwise we could remove NA with:
auto <- 
  Auto %>%
  drop_na()
```

Quantitative or qualitative predictors?

All predictors are quantitative besides `name`.
`cylinders` is technically numerical, but given is strict discrete nature, could it be considered categorical?

```{r}
auto %>%
  map(class) %>%
  # it looks better as a named character,
  # but remember the curly braces
  {set_names(flatten_chr(.),
             names(.))}
```

What is the range of the quantitative predictors?

```{r}
auto %>%
  select_if(is.numeric) %>%
  # map_if(.x = .,
  #        .p = is.numeric,
  #        .f = range)
  map(range) %>%
  # facultative
  # tibbles print better than lists
  as_tibble()
```

And what is their mean and standard dev?

```{r}
auto %>%
  select_if(is.numeric) %>%
  # map the mean and sd function to each numeric column
  # and directly stores the results in a tibble
  {tibble(var = colnames(.),
          mean = map_dbl(., mean),
          sd = map_dbl(., sd))} %>%
  # visualize in markdown / HTML
  knitr::kable()
```

Remove 10th to 85th observation and check again.

```{r}
auto %>%
  # this was easier to do without dyplr
  .[-c(10:85), ] %>% 
  # the rest is the same
  select_if(is.numeric) %>%
  {tibble(var = colnames(.),
          mean = map_dbl(., mean),
          sd = map_dbl(., sd))} %>%
  # visualize in markdown / HTML
  knitr::kable()
```

Explore the distribution and the relation among the variables

```{r, results='hide', message=FALSE}
auto %>% 
  select_if(is.numeric) %>%
  # ggpairs from the package GGally 
  # is a good alternative to pairs
  GGally::ggpairs()
```

How did horsepower changed during the years?

```{r}
library(lubridate)

p <- 
auto %>%
  ggplot(aes(x = year %>% as.character() %>% as_factor(),
             y = horsepower))

p + geom_boxplot() + geom_point()

p + geom_boxplot() + geom_point(aes(colour = weight))

p2 <- 
  auto %>% 
  ggplot(aes(x = year %>% as.character() %>% as_factor(),
             y = horsepower/weight))

p3 <- p2 + 
  geom_boxplot() + 
  ggbeeswarm::geom_quasirandom()

p3 + 
  ggrepel::geom_text_repel(data = auto %>% 
                             # mutate(var = horsepower/weight) %>%
                             filter(horsepower/weight > 0.052),
                           aes(label = name)) +
  theme_bw()
```

Which variable could we use to predict Gas Mileage (mpg)?

From the paired scatterplot, weight might be the best predictor, but also horsepower and displacement. They all correlate negatively with mpg.

```{r}

auto %>%
  select(name, mpg, displacement, horsepower, weight) %>%
  # a bit unorthodox way to facet scatterplot
  gather(key = variable,
         value = value,
         displacement,
         horsepower,
         weight) %>%
  ggplot(aes(x = value, 
             y = mpg)) +
  geom_point(alpha = .3, colour = "grey30") + 
  facet_wrap(facets = "variable", 
             scales = "free_x") +
  theme_bw()
```

## Explore the Boston Housing Dataset [10]

We can access the `Boston` dataset from the package `MASS` without attaching the package itself with `::` as in: `MASS::Boston`.

```{r}

# tibbles print better
boston <- 
  MASS::Boston %>%
  as_tibble()

# indeed the tibble print method 
# displays line and row number automatically
boston
```

Make pairwise scatterplots, describe your findings

```{r, results='hide', message=FALSE, fig.width=30, fig.retina=TRUE}
boston %>% 
  GGally::ggpairs()
```

some observations:

- Criminality has a spike in most correlations plots, like if it was concentrated in one area only.

- Nox increases with Industry and Age, and decreases with the distance to the employment centers.

- Tax decreases with median value of user occupied homes?



